---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---
```{r global_options, include=F}
library(knitr)
opts_chunk$set(cache=T, message=F, warning=F, error=T)
```

## Loading and preprocessing the data
First let's read the zipped data:
``` {r read_data}
data <- read.csv(unzip("activity.zip"), 
                 na.strings = "NA", 
                 colClasses = c("integer", "Date", "integer")
                 )
```
Let's see what we've got:
``` {r data_summary}
summary(data)
head(data)
```


## What is mean total number of steps taken per day?
We begin by grouping the data by day and finding the total steps per day:
``` {r daily_steps}
library(dplyr)
by_day <- group_by(data, date)
daily_steps <- summarize(by_day, total_steps = sum(steps, na.rm=T))
```

Now we can see the distribution of the daily total steps:
``` {r histogram}
hist(daily_steps$total_steps, 
     breaks=20,
     xlab="mean daily steps", main=""
     )
```

To get a better sense of the data we can also compute the: 

* mean: **`r mean(daily_steps$total_steps)`** with ` mean(daily_steps$total_steps)`  
* median: **`r median(daily_steps$total_steps)`** with ` median(daily_steps$total_steps)`


## What is the average daily activity pattern?
Similarly, we can see the average number of steps grouped by the 5-minute time interval of a day, over all the days in our data set:

``` {r interval_steps}
by_interval <- group_by(data, interval)
interval_steps <- summarize(by_interval, avg_steps = mean(steps, na.rm=T))
```

Plotting this, we have:
``` {r interval_plot}
library(lattice)
xyplot(interval_steps$avg_steps ~ interval_steps$interval, 
       data=interval_steps, 
       type="l",
       xlab="5-minute interval of day",
       ylab="average number of steps"
       )
```
  
It might be interesting to know which interval has the highest average number of steps. We can find this easily:
``` {r max_interval}
interval_steps[which.max(interval_steps$avg_steps), "interval"]
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
