---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---
```{r global_options, include=F}
library(knitr)
opts_chunk$set(cache=F, message=F, warning=F, error=T)
options(scipen = 2, digits = 2)
```

## Loading and preprocessing the data
First let's read the zipped data:
``` {r read_data}
data <- read.csv(unzip("activity.zip"), 
                 na.strings = "NA", 
                 colClasses = c("integer", "Date", "integer")
                 )
```
Let's see what we've got:
``` {r data_summary}
summary(data)
head(data)
```


## What is mean total number of steps taken per day?
We begin by grouping the data by day and finding the total steps per day:
``` {r daily_steps}
library(dplyr)
by_day <- group_by(data, date)
daily_steps <- summarize(by_day, total_steps = sum(steps))
```

Now we can see the distribution of the daily total steps:
``` {r histogram}
hist(daily_steps$total_steps, 
     breaks=20,
     xlab="mean daily steps", 
     main="",
     col="skyblue"
     )

abline(v = mean(daily_steps$total_steps, na.rm = T), col = "red")
```

To get a better sense of the data we can also compute the: 

* mean: **`r mean(daily_steps$total_steps, na.rm = T)`** with ` mean(daily_steps$total_steps, na.rm = T)` 

* median: **`r median(daily_steps$total_steps, na.rm = T)`** with ` median(daily_steps$total_steps, na.rm = T)`


## What is the average daily activity pattern?
Similarly, we can see the average number of steps grouped by the 5-minute time interval of a day, over all the days in our data set:

``` {r interval_steps}
by_interval <- group_by(data, interval)
interval_steps <- summarize(by_interval, avg_steps = mean(steps, na.rm=T))
```

Plotting this, we have:
``` {r interval_plot}
library(lattice)
xyplot(interval_steps$avg_steps ~ interval_steps$interval, 
       data=interval_steps, 
       type="l",
       xlab="5-minute interval of day",
       ylab="average number of steps"
       )
```
  
It might be interesting to know which interval has the highest average number of steps. We can find this easily:
``` {r max_interval}
interval_steps[[which.max(interval_steps$avg_steps), "interval"]]
```


## Imputing missing values
Earlier we saw that there are **`r sum(is.na(data))`** missing, or `NA`, values. This can be calculated using `sum(is.na(data))`.  

Now we might wish to see the impact these missing values have. Let's replace `NA`s with the average (mean) number of steps for that interval.  

``` {r fill_NAs}
fdata <- left_join(data, interval_steps) 
fdata$steps <- ifelse(is.na(fdata$steps), fdata$avg_steps, fdata$steps)
```

As before, we'll plot a histogram of the total steps taken:

``` {r filled_hist}
f_daily_steps <- summarize(group_by(fdata, date),
                           total_steps = sum(steps)
                           )

hist(f_daily_steps$total_steps, 
     breaks=20,
     xlab="mean daily steps", 
     main="",
     col="skyblue"
     )

abline(v = mean(f_daily_steps$total_steps, na.rm = T), col = "red")
```

As we might have expected, the counts have increased compared to the original histogram, since the `NA`s, which were previously not included, have now been assigned average values. The mean and median are now identical, suggesting the data are no longer skewed.

* mean: **`r mean(f_daily_steps$total_steps)`** with ` mean(f_daily_steps$total_steps)`  
* median: **`r median(f_daily_steps$total_steps)`** with ` median(f_daily_steps$total_steps)`


## Are there differences in activity patterns between weekdays and weekends?
To compare activity levels between weekdays and weekends, we'll create a column of factor variables identifying the time of the week.
``` {r time_of_week}
fdata$tow <- as.factor(ifelse(weekdays(fdata$date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))

```

Now, we can compute the average number of steps per interval as before, but this time we'll do it separately for weekdays and weekends.

``` {r time_of_week_avg_steps}
tow_steps <- summarize(group_by(fdata, interval, tow),
                       avg_steps = mean(steps, na.rm=T)
                       )
```

Finally, let's plot the results to compare weekday and weekend activity:

``` {r comp_plot}
xyplot(avg_steps ~ interval | tow,
       data=tow_steps, 
       type="l",
       xlab="5-minute interval of day",
       ylab="average number of steps",
       layout=c(1,2)
       )

```


